Acceleration of Bilateral filtering algorithm for manycore and multicore architectures

### 2012-03-10 LateX/paper8.tex
test
editing wording of the paragraph about divergent branching. e.g. changed will to would etc

CUDA's intra-warp implicit synchronization capability is the cornerstone of our CUDA implementation of our pair-symmetric algorithm. Without implicit synchronization, our pair-symmetric algorithm would suffer significantly over CUDA because it would require expensive synchronization calls per Center-Neighbor Filter computation.

In CUDA, threads are organized into groups called warps. Each warp executes one instruction at a time for its own group of threads. Because all threads in a warp share the same program counter and are therefore driven by the same sequential instruction stream, all threads in a warp are always in sync with each other without requiring explicit synchronization calls. Therefore, synchronization between threads in a warp is implicit.

To take full advantage of CUDA's implicit intra-warp synchronization feature, divergent branching between threads in the same warp must be avoided. This is so because all the threads in the same warp share only one program counter. A divergent branch will cause some or most of the threads to be idle as the program counter runs through the branch and drives only those threads that meet the branch condition to work. An example of a divergent branch can be a simple if-statement such as: $if~(threadIdx.x < 5)~\{...\}~else~\{...\}$.

Divergent branching will commonly occur in memory access patterns in which global memory operations are not coalesced and shared memory operations are not free of memory bank conflicts. For example, if eight threads were to read from eight non-contiguous segments of global memory - commonly seen in strided access patterns like array {[}threadIdx.x {*} 2{]} - divergent branching would occur because a warp of threads can only read from one segment of memory per instructional transaction, and that segment's word-length is hardware defined.

In other words, a memory transaction will be serialized unless all the threads in the warp to which the transaction belongs are aligned consecutively and fit within a segment of memory. Therefore, in order for all the threads of a warp to read/write from global memory in a single transaction, the threads must correspond read/write operations to aligned contiguous addresses within a segment of memory. For Bilateral filtering algorithm, warps should correspond to horizontal rows of pixels for an [i*jMax + j] representation of a 2D image, for if threads in a warp were to operate on pixels in a vertical order, there would be a separate transaction for each thread in the warp.

>>>>

CUDA's intra-warp implicit synchronization capability is the cornerstone of our CUDA implementation of our pair-symmetric algorithm. Without implicit synchronization, the performance of our CUDA-implemented pair-symmetric algorithm would be significantly reduced because it would require expensive synchronization calls per Center-Neighbor Filter computation.

In CUDA, threads are organized into groups called warps, and warps are launched in groups called thread-blocks, in which each thread-block can share among its warps an allocated portion of shared-memory. Each warp executes one instruction at a time for its own group of threads. Because all threads in a warp share the same program counter and are therefore driven by the same sequential instruction stream, all threads in a warp are always in sync with each other without requiring explicit synchronization calls. Hence, synchronization between threads in a warp is implicit.

To take full advantage of CUDA's implicit intra-warp synchronization, divergent branching between threads in the same warp must be avoided. This is because all the threads in the same warp share only one program counter. A divergent branch would cause some or most of the threads to be idle as the program counter would run through the branch and drive only those threads that meet the branch condition to work. An example of a divergent branch can be a simple if-statement such as: $if~(threadIdx.x < 5)~\{...\}~else~\{...\}$.

Divergent branching commonly occurs in memory access patterns in which global memory operations are not coalesced and shared memory operations are not free of memory bank conflicts. For example, if eight threads were to read from eight non-contiguous segments of global memory---commonly seen in strided access patterns like array {[}threadIdx.x {*} 2{]}---divergent branching would occur because a warp of threads can only read from one segment of memory per instructional transaction, and that segment's word-length is hardware defined. In other words, a memory transaction would be serialized if all the threads in the warp to which the transaction belonged were not aligned consecutively or did not fit within a segment of memory. Therefore, in order for all the threads of a warp to read/write from global memory in a single transaction, the threads must correspond read/write operations to aligned contiguous addresses within a segment of memory. For the Bilateral filtering algorithm, warps should correspond to horizontal rows of pixels for an [i*j_pitch + j] representation of a 2D image, for if threads in a warp were to operate on pixels in a vertical order, there would be a separate transaction for each thread in the warp.
